---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# WHcircular

<!-- badges: start -->
<!-- badges: end -->

This package re-implements circular-linear regression and other selected functionality from the `circular` package for improved performance, particularly by avoiding construction of $O(n^2)$ diagonal matrices. Additional functions are provided for time-of-day variables in "HH:MM" format. 'WH' code is designed to maximize speed, for use in simulation studies. As such, input validation is minimal and the user is responsible for ensuring valid use.

## Installation

You can install the development version of WHcircular from [GitHub](https://github.com/hongconsulting/WHcircular) with:

``` r
remotes::install_github("hongconsulting/WHcircular")
```

## Example 1

```{r example1}
library(WHcircular)
y <- c("01:00","01:15","01:30","01:45","02:00",
       "23:00","23:15","23:30","23:45","00:00")
X <- c(rep(0, 5), rep(1, 5))
print(HHMM.mean(y))
print(HHMM.lm(y, ~X))
```

## Example 2

```{r example2}
library(WHcircular)
library(nycflights13)
data <- nycflights13::flights
y <- sprintf("%04d", data$sched_dep_time)
y <- paste0(substr(y, 1, 2), ":", substr(y, 3, 4))
print(HHMM.lm(y = y, X = ~as.factor(data$origin),
              name = c("EWR", "JFK", "LGA")))
```

## Example 3

```{r example3}
library(WHcircular)
y <- c("01:00","01:15","01:30","01:45","02:00",
       "23:00","23:15","23:30","23:45","00:00")
X <- c(rep(0, 5), rep(1, 5))
theta <- WH_HHMM_to_rad(y)
print(WH_rad_to_HHMM(WH_rad_mean(theta)))
print(WH_rad_to_HHMM(WH_rad_sd(theta)))
fit <- WH_reg_circlinear(theta, as.matrix(X))
print(fit$coef)
print(fit$SE)
print(WH_rad_to_HHMM((fit$coef[2])))
print(WH_rad_to_dHHMM(2 * atan(fit$coef[3])))
print(WH_CordeiroPaulaBotter(theta, X))
```

## Further Reading

1. Cordeiro, G.M., Paula, G.A. and Botter, D.A., 1994. Improved likelihood ratio tests for dispersion models. *International Statistical Review*, pp. 257–274.
2. Fisher, N.I. and Lee, A.J., 1992. Regression models for an angular response. *Biometrics*, pp. 665–677.
